---
title: "H20: Generalized Linear Models (GLM)"
author: '@jrcajide'
date: "8/4/2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval=FALSE, include=FALSE}
# The following two commands remove any previously installed H2O packages for R.
if ("package:h2o" %in% search()) { detach("package:h2o", unload=TRUE) }
if ("h2o" %in% rownames(installed.packages())) { remove.packages("h2o") }
```

```{r, eval=FALSE, include=FALSE}
# Next, we download packages that H2O depends on.
if (! ("methods" %in% rownames(installed.packages()))) { install.packages("methods") }
if (! ("statmod" %in% rownames(installed.packages()))) { install.packages("statmod") }
if (! ("stats" %in% rownames(installed.packages()))) { install.packages("stats") }
if (! ("graphics" %in% rownames(installed.packages()))) { install.packages("graphics") }
if (! ("RCurl" %in% rownames(installed.packages()))) { install.packages("RCurl") }
if (! ("jsonlite" %in% rownames(installed.packages()))) { install.packages("jsonlite") }
if (! ("tools" %in% rownames(installed.packages()))) { install.packages("tools") }
if (! ("utils" %in% rownames(installed.packages()))) { install.packages("utils") }

# Now we download, install and initialize the H2O package for R.
install.packages("h2o", type="source", repos=(c("http://h2o-release.s3.amazonaws.com/h2o/rel-slater/5/R")))
```


```{r}
# install.packages("h2o")
library(h2o)

h2o.init(nthreads = 4, max_mem_size = "2G")

# h2o.clusterInfo()
```

Generalized Linear Models (GLM): Provides flexible generalization of
ordinary linear regression for response variables with error distribution models
other than a Gaussian (normal) distribution. GLM unifies various other
statistical models, including Poisson, linear, logistic, and others when using `1
and `2 regularization.

```{r}
# Import dataset and display summary
airlinesURL = "https://s3.amazonaws.com/h2o-airlines-unpacked/allyears2k.csv"
airlines.hex = h2o.importFile(path = airlinesURL, destination_frame = "airlines.hex")
summary(airlines.hex)
```

```{r}
# View quantiles and histograms
#high_na_columns = h2o.ignoreColumns(data = airlines.hex)
quantile(x = airlines.hex$ArrDelay, na.rm = TRUE)
```

```{r histogram}
h2o.hist(airlines.hex$ArrDelay)
```


#### Find number of flights by airport

```{r}
originFlights = h2o.group_by(data = airlines.hex, by = "Origin", nrow("Origin"), gb.control=list(na.methods="rm"))
originFlights.R = as.data.frame(originFlights)
knitr::kable(head(originFlights.R))
```

#### Find number of flights per month

```{r}
flightsByMonth = h2o.group_by(data = airlines.hex, by = "Month", nrow("Month"),gb.control=list(na.methods="rm"))
flightsByMonth.R = as.data.frame(flightsByMonth)
knitr::kable(head(flightsByMonth.R))
```

#### Find months with the highest cancellation ratio

```{r}
which(colnames(airlines.hex)=="Cancelled")
cancellationsByMonth = h2o.group_by(data = airlines.hex, by = "Month", sum("Cancelled"),gb.control=list(na.methods="rm"))
cancellation_rate = cancellationsByMonth$sum_Cancelled/flightsByMonth$nrow
rates_table = h2o.cbind(flightsByMonth$Month, cancellation_rate)
rates_table.R = as.data.frame(rates_table)
knitr::kable(head(rates_table.R))
```

#### Construct test and train sets using sampling

```{r}

airlines.split = h2o.splitFrame(data = airlines.hex,ratios = 0.85)
airlines.train = airlines.split[[1]]
airlines.test = airlines.split[[2]]

head(airlines.train)
```

#### Display a summary using table-like functions

```{r}
h2o.table(airlines.train$Cancelled)
h2o.table(airlines.test$Cancelled)
```

#### Set predictor and response variables

```{r}
Y = "IsDepDelayed"
X = c("Origin", "Dest", "DayofMonth", "Year", "UniqueCarrier", "DayOfWeek", "Month", "DepTime", "ArrTime", "Distance")
```

#### Define the data for the model and display the results

```{r}
airlines.glm <- h2o.glm(training_frame=airlines.train, x=X, y=Y, family = "binomial", alpha = 0.5)
```

#### View model information: training statistics, performance, important variables
```{r}
summary(airlines.glm)
```

#### Coefficients

```{r}
h2o.coef(airlines.glm)
```


#### Variable Importances

```{r}
h2o.varimp(airlines.glm)
```

#### Predict using GLM model

```{r}
pred = h2o.predict(object = airlines.glm, newdata = airlines.test)
```

#### Look at summary of predictions: probability of TRUE class (p1)

```{r}
summary(pred)
```

